---
title: "BIA 6301 APPLIED DATA MINING HOMEWORK ASSIGNMENT 1"
author: "Leonardo Ji"
date: '`r Sys.Date()`'

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

##Part A. Underwriting at KC Health 

The purpose of this assignment is to simulate underwriting at a health insurance company called KC Health.

You were recently hired as a data scientist in the underwriting team at KC Health. The Vice President of Underwriting, who is a Fellow of the Society of Actuaries, gave you two data files during your first week on the job. The first data file is named Insurance.csv, which contains last yearss medical charges of 1,338 customers. The data file also contains information of six demographic and geographic variables: age, sex, BMI, children, smoker, and region. 

### Data

- **Age**: age of the customer
- **Sex**: sex of the customer
- **BMI**: reported body mass index (BMI) of customer
- **Children**: number of children reported by the customer
- **Smoker**: whether the customer is a smoker
- **Region**: region of the country where the customer resides. Four regions are included: northeast, northwest, southeast, and southwest. 


```{r}
insuranceData<-read.csv(file="C:/Users/lj015625/Desktop/DataMining Class/Applied_Data_Mining/assignment_module_1/data/insurance.csv")
str(insuranceData)
summary(insuranceData)
dim(insuranceData)
names(insuranceData)
```


```{r}
insuranceData$obesity_category <- ifelse(insuranceData$bmi >= 30, 'yes', 'no')
insuranceData$obesity <- ifelse(insuranceData$bmi >= 30, 1, 0)

insuranceData$smokerFactor <- ifelse(insuranceData$smoker=='yes', 1, 0)

insuranceData$obesity_smoker <- insuranceData$obesity*insuranceData$smokerFactor
insuranceData$obesity_smoker_category <- ifelse(insuranceData$obesity_smoker == 1, 'yes', 'no')

library(ggplot2)
ggplot(insuranceData, aes(x = age, y = charges)) +
  geom_point(color = "forestgreen", shape=2, size = 1) + theme_bw() + 
  stat_smooth(method = "lm", col = "red", se=FALSE) +
  xlab("Customer Age") + ylab("Charges in USD") +
  ggtitle("Customer Age vs. Charges") 

ggplot(insuranceData, aes(x = children, y = charges)) +
  geom_point(color = "blue", shape=1, size = 1) + theme_bw() + 
  stat_smooth(method = "lm", col = "red", se=FALSE) +
  xlab("Children") + ylab("Charges in USD") +
  ggtitle("Children vs. Charges") 

ggplot(insuranceData, aes(fill = obesity_category, factor(obesity_category), charges)) + 
  geom_boxplot() + theme_bw() + xlab("Obesity") + ylab("Charges in USD") +
  ggtitle("Obesity vs non-obesity Charges") 

ggplot(insuranceData, aes(fill = smoker, factor(smoker), charges)) +
  geom_boxplot() + theme_bw() + xlab("Sex") + ylab("Charges in USD") +
  ggtitle("Non-smoker vs. Smoker Charges")

ggplot(insuranceData, aes(fill = obesity_smoker_category, factor(obesity_smoker_category), charges)) +
  geom_boxplot() + theme_bw() + xlab("Obesity and Smoker") + ylab("Charges in USD") +
  ggtitle("Normal vs. Obesity and Smoker Charges")
```


###The VP has six questions for you to answer:

####1.	Which demographic and geographic variables explain medical charges of the 1,338 customers in the insurance.csv file? How do you know? What are the effects of these identified variables on medical charges? 

Age, bmi, children, smoker.  Those variables all have very small p-value (99.9% confidence). 


```{r}
insuranceFullModel<-lm(charges~.,data=insuranceData)
summary(insuranceFullModel)
```

If we remove insignificant terms sex and region.

$Charges = -12102.77 + 257.85 *age + 321.85 * bmi + 473.50 * children + 23811.40 * smoker$

```{r}
reducedInsuranceModel <- lm(charges~ age+bmi+children+smoker,data=insuranceData)
summary(reducedInsuranceModel)
```


####2.	The World Health Organization (WHO) defined an obese person as someone who has a BMI ≥ 30. What is the effect of obesity on medical charges? [Hint: You will need to use the specified model in Question 1 and add in a dummy variable to account for the effect of obesity.]

Obesity (bmi >= 30) increases medical charges by 2904.47.

$Charges = -7816.28 + 258.03 *age + 131.71 * bmi + 473.87 * children + 23819.41 * smoker + 2904.47 * obesity$

```{r}
insuranceData$obesity <- (ifelse(insuranceData$bmi >= 30, 1, 0))
insuranceModel2 <- lm(charges~age+bmi+children+smoker+obesity, data=insuranceData)
summary(insuranceModel2)
```

####3.	What is the effect on medical charges of someone who is both obese and a smoker? [Hint: You will need to use the specified model in Question 2 and add in an interaction term to account for obesity and smoking.]

Obseity and smoker increases medical charges by 19684.869.

$Charges = -5097.881 + 264.226 *age + 97.698 * bmi + 512.424 * children + 512.424 * smoker - 809.791 * obesity + 19684.869 * obesity smoker$

```{r}
insuranceData$smokerFactor <- (ifelse(insuranceData$smoker=='yes', 1, 0))
insuranceData$obesity_smoker <- (insuranceData$obesity*insuranceData$smokerFactor)
library(knitr)
Numeric_Vars <-insuranceData[,c('age','bmi','children','smokerFactor','obesity_smoker','charges')]
Correlation_Matrix <- cor(Numeric_Vars)
kable(Correlation_Matrix)
```
```{r}
insuranceModel3 <- lm(charges~age+bmi+children+smoker+obesity+obesity_smoker, data=insuranceData)
summary(insuranceModel3)

```

####4.	Can you recommend a preferred model Health KC should use to predict medical charges of potential customers? How much variation in the Insurance dataset is explained by the preferred model? [Hint: You will need to conduct partial F-tests to compare the three models created to answer Questions 1, 2, and 3.]

The anova() function performs a hypothesis test comparing the three models. 
First we compare reduced model with obesity modelgg. The null hypothesis is that the two models fit the data equally well, and the alternative hypothesis is that the model with obesity is superior.  F-statistic is 28.16, p- value is very small.  So we reject null hypothesis.
Next we compare model with obesity with model with obesity smoker.  The null hypothesis is that the two models fit the data equally well, and the alternative hypothesis is that the model with obesity smoker is superior.  F-statistic is 21004650148, p-value is very small.  So we reject null hypothesis.

We can say the model three with obesity smoker is the preferred model.

```{r}
anova(reducedInsuranceModel, insuranceModel2)
anova(insuranceModel2, insuranceModel3)
```

####5.	Are there any known statistical problem(s) with the chosen model? [Hint: You will need to examine the four linear regression assumptions plus multicollinearity.] If there are known problem(s), what are your recommendation(s) for mitigating them?

Residual vs Fitted Plot-The plotted values should look random (i.e. no pattern). If the plot shows a curvilinear relationship, then assumption #4 (linearity) is violated. The residuals is mostly positive.  This means model is underestimate true predicted value.  

Normal Q-Q Plot-The plotted values should be a straight line. If it is curved (an S shape), then assumption #2 (normality of errors) is violated.
This means data has more extreme value than would be expected if they truly came from a Normal distribution.  The higher medical charges values is not a truely normal distribution.

Scale-Location Plot-This plot is a repeat of the Residuals vs. Fitted Plot but on a different scale. The square root of the standardized residuals are plotted against the fitted values. The standardized residuals should look random. If there is a pattern the values are most positive assumption #3 (homoskedasticity) is violated.

Residuals vs. Leverage-This plot highlights the outlier values of the target variable (Y) that have the highest effect on the parameter estimates. These are called “high influence” points. Look for values outside of the Cook’s distance (red dashed line).  There are many outliers above the Cook's distance line.  There are one single outlier far below Cook's distance line.

The recommendations are to remove outlier charges, gather three times more rows from vif test.  We also need more variables that can explain the higher charges.


```{r}
library(ggfortify)
autoplot(insuranceModel3, label.size = 3) 
```

Multicollinearity
Use variance inflation factor (VIF) for each predictors
Here obesity is close to 3 which means some autocorrelation.  We need a sample size three times as large as your current sample size to overcome the multicollinearity in the model.


```{r}
library(car)
vif(insuranceModel3)

```
No Autocorrelation
Use Durbin Watson statistic. DW statistic close to 2 means there is no autocorrelation. Closer to 0 means positive autocorrelation. Closer to 4 means negative autocorrelation.
Here value is 2.05 so we can say there is no autocorrelation.


```{r}
durbinWatsonTest(insuranceModel3)
```

Homoskedasticity
The probability distribution of the errors has constant variance.
P-value is slightly above 0.05 so we cannot rule out heteroscedasticity.

```{r}
ncvTest(insuranceModel3)
```


####6.	The VP gives you a batch file called CompanyABC.csv, which contains demographic information of 50 potential customers at a small business. All 50 employees are scattered across the eastern U.S. seaboard. Use the preferred model to predict medical charges of the 50 potential customers in the CompanyABC.csv batch file. She does not want to see predicted charges for all 50 customers. The VP only wants the predictions summarized at a high level for ease of communications with other managers at Health KC. [Hint: You should create a new variable in your data frame for the predicted charges. Something simple like “estimated_charges” is sufficient. You will then perform exploratory data analysis of “estimated_charges” to examine the 50 predicted charges.]

The estimated mean charge is 12181 per customer. The estimate median charge is 10436.
The estimated minimum charge is 1791. The estimated maximum charge is 47264.  


```{r}
testingData<-read.csv(file="C:/Users/lj015625/Desktop/DataMining Class/Applied_Data_Mining/assignment_module_1/data/CompanyABC.csv")
testingData$obesity <- ifelse(testingData$bmi >= 30, 1, 0)
testingData$obesity_category <- ifelse(testingData$bmi >= 30, 'yes', 'no')
testingData$smokerFactor <- ifelse(testingData$smoker=='yes', 1, 0)
testingData$obesity_smoker <- testingData$obesity*testingData$smokerFactor
testingData$obesity_smoker_category <- ifelse(testingData$obesity_smoker == 1, 'yes', 'no')
summary(testingData)

testingData$estimated_charges <- predict(insuranceModel3, testingData)
summary(testingData$estimated_charges)

```

A smoker or an obesity customer have higher estimate charges.
Both smoker and obesity customer have far higher estimate charges due to interaction term from smoker and obesity.

```{r}
aggregate(estimated_charges~smoker, data=testingData, FUN="mean")
aggregate(estimated_charges~obesity, data=testingData, FUN="mean")
# interaction term
aggregate(estimated_charges~obesity_smoker, data=testingData, FUN="mean")
```


ggplot helps visualize data much better than tables.

```{r}
library(ggplot2)
ggplot(testingData, aes(x = age, y = estimated_charges)) +
  geom_point(color = "forestgreen", shape=4, size = 2) + theme_bw() + 
  stat_smooth(method = "lm", col = "red", se=FALSE) +
  xlab("Customer Age") + ylab("Estimated Charges in USD") +
  ggtitle("Customer Age vs. Estimate Charges") 

ggplot(testingData, aes(x = children, y = estimated_charges)) +
  geom_point(color = "blue", shape=1, size = 2) + theme_bw() + 
  stat_smooth(method = "lm", col = "red", se=FALSE) +
  xlab("Children") + ylab("Estimated Charges in USD") +
  ggtitle("Children vs. Estimate Charges") 

ggplot(testingData, aes(fill = obesity_category, factor(obesity_category), estimated_charges)) + 
  geom_boxplot() + theme_bw() + xlab("Obesity") + ylab("Estimated Charges in USD") +
  ggtitle("Obesity vs non-obesity Estimate Charges") 

ggplot(testingData, aes(fill = smoker, factor(smoker), estimated_charges)) +
  geom_boxplot() + theme_bw() + xlab("Sex") + ylab("Estimated Charges in USD") +
  ggtitle("Non-smoker vs. Smoker Estimate Charges")

ggplot(testingData, aes(fill = obesity_smoker_category, factor(obesity_smoker_category), estimated_charges)) +
  geom_boxplot() + theme_bw() + xlab("Obesity and Smoker") + ylab("Estimated Charges in USD") +
  ggtitle("Normal vs. Obesity and Smoker Estimate Charges")
```







End

